Performance-portable Parallel Programming in Computer Science Education with Kokkos

### 2020-01-23 src/4.ProgammingModel.tex
Added parallel_for example.

Figure~\ref{fig:parallelFor} shows the definition of the Kokkos~\emph{parallel\_for} interface. The interface accepts two template parameters and three function arguments. \emph{ExecPolicy} shapes the iteration space for a particular execution space. Valid template parameter are the~\emph{RangePolicy},~\emph{TeamPolicy},~\emph{MDRangePolicy} and others. As the name suggests, these policies express how an iteration space is mapped to the underlying hardware. In particular, a range policy defines a one dimensional iteration range. Team policies allow to map an iteration space to architectures that expose hierarchical parallelism.. Policies commonly accept template parameters such as~\emph{execution space},~\emph{schedule} or~\emph{index type}. For the interested reader, the complete programming model specification can be accessed on-line\cite{pub:KOKKOS}.

Programming with patterns and abstractions allows the compiler to generate optimized transformations. In case of parallel patterns, the execution is unspecified and only promises deterministic results for the reduction and scan operations. This gives the freedom to the underlying abstraction layers to implement different mapping patterns on different hardware such as assignment of iterations to threads or vector lanes. We believe that these abstraction and their descriptive semantic represent key programming primitives and the way forward for native language support for parallel programming.

Figure~\ref{fig:SemCaptureKokkos} shows the semantic capture defined by the Kokkos programming model. 
%Figure~\ref{fig:stack} shows abstraction layers in Kokkos and and how it fits in a parallel programming landscape.
The next section shows how the Kokkos parallel\_for class is mapped to the the underlying back-end programming model.

@@

>>>>

Figure~\ref{fig:parallelFor} shows one definition of the Kokkos~\emph{parallel\_for} interface. This interfac accepts two template parameters and three function arguments. \emph{ExecPolicy} shapes the iteration space for a particular execution space. Valid template parameter are the~\emph{RangePolicy},~\emph{TeamPolicy},~\emph{MDRangePolicy},~\emph{TeamThreadRange} and~\{ThreadVectorRange}. As the name suggests, these policies express how an iteration space is mapped to the underlying hardware. For example, a range policy defines a one dimensional iteration range. A team policy defines a 1D iteration range, each of which is assigned to a thread team. This policy allows the expression of hierarchical parallelism. For the interested reader, the complete programming model specification can be accessed on-line\cite{pub:KOKKOS}. Figure~\ref{fig:abstractions} lists patterns and abstractions and structures them by information type in the semantic capture.

Programming with patterns and abstractions allows the compiler to generate optimized transformations. In case of parallel patterns, the execution is unspecified and only promises deterministic results for the reduction and scan operations. This gives the freedom to the underlying abstraction layers to implement different mapping patterns on different hardware such as assignment of iterations to threads or vector lanes. We believe that these abstraction and their descriptive semantic represent key programming primitives and the way forward for generic support for parallel programming. In conclusion, Figure~\ref{fig:SemCaptureKokkos} shows the semantic capture defined by the Kokkos programming model. 
%Figure~\ref{fig:stack} shows abstraction layers in Kokkos and and how it fits in a parallel programming landscape.
The next section shows how the Kokkos parallel\_for class is mapped to the the underlying back-end programming model.

@@

\begin{figure}
\centerline{\includegraphics[width=0.3\textwidth]{img/Abstractions.png}}
\caption{Landscape}
\label{fig:abstractions}
\end{figure}
